---
author: "Brian Kreis"
date: "March 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width=12, fig.height=8)

library(knitr)
library(broom)
```

## Model 4 - Base Runs (BsR)
\
In this model we incorporate our calculated metric Base Runs(BsR), a sabermetric stat created by David Smyth, to predict the number of runs a team would be expected to have scored based on the types of hits and number of walks. We include all available variables, beginning with BsR, and use forward stepwise regression to add statistically significant variables to the model.

BsR is calulated as follows:

\
$$\frac{A * B}{B + C} + D \\ 
\\
\\
$$

$$ A  = Hits + Walks - Home Runs \\
B = (1.4*Total Bases - 0.6*Hits - 3*HomeRuns + 0.1*Walks) \\
C = AB - H \\
D = HR \\
\\
$$

\
We approximated the average at bats for a team in a 162 game season infromation from baseball reference. 
https://www.baseball-reference.com/leagues/MLB/bat.shtml

```{r}

data <- read.csv("https://raw.githubusercontent.com/vbriot28/Data621_group2/master/data_group2_nbc.csv")

model4 <- step(lm(TARGET_WINS ~ BsR, data = data),
                  direction = "forward", 
               scope = ~ BsR + BATTING_2B + BATTING_3B + BATTING_HR + BATTING_BB + BATTING_SO + 
                 BASERUN_SB + PITCHING_H + PITCHING_HR + PITCHING_BB + PITCHING_SO + 
                 FIELDING_E + FIELDING_DP + BATTING_1B + BATTING_TB + WHGP + PITCHING_SO_BB + BATTING_BB_SO 
)


tbl4 <- tidy(model4)
kable(tbl4)

par(mfrow=c(2,2)) 
plot(model4)


```

## Model 4B

As variables were added to the model, the statistical significance of some initial variables was reduced. In fact our main statistic of interest, BsR, is no longer statistically significant. In order to develop the best selection of variables, we also incorporate a bidrectional method to revisit the significance of variables added earlier in the analysis.


```{r}

model4b <- step(lm(TARGET_WINS ~ BsR, data = data),
               direction = "both", 
               scope = ~ BsR + BATTING_2B + BATTING_3B + BATTING_HR + BATTING_BB + BATTING_SO + 
                 BASERUN_SB + PITCHING_H + PITCHING_HR + PITCHING_BB + PITCHING_SO + 
                 FIELDING_E + FIELDING_DP + BATTING_1B + BATTING_TB + WHGP + PITCHING_SO_BB + BATTING_BB_SO 
)


tbl4b <- tidy(model4)
kable(tbl4b)


par(mfrow=c(2,2)) 
plot(model4b)


```

Using the bidirectional approach, BsR was removed from the model. Once BsR was removed, BATTING_2B, BATTING_1B and BATTING_3B regained their significance. This is ikely caused by collinearity within the variables as BsR is a derived stat based in large part on hits. Because BsR was found to not add predictive ability to our model, Model 4B is the superior model.





